---
title: 'Group 3 - Chapter 9: Support Vector Machines Tutorial'
author:
- Jolien Lam (j.c.lam@amc.uva.nl) – University of Amsterdam - jolienlam
- Yeng Moua (moua0199@umn.edu) – University of Minnesota -
- Fortuna Vuthaj (fortuna.vuthaj@gmail.com) – University of Heilbronn/Heidelberg
  -
output:
  html_document:
    df_print: paged
---

```{r setup, include= FALSE}


`%!in%` = Negate(`%in%`)
```



This notebook documents work by group 3 utilizing support vector machines in the analysis of health data. Informaiton about the project can be found [here](https://iphie2018.wordpress.com/academic-program/).

# Front Matter

# Extraction

Data were acquired from ....

Two files are retrievable. One of the aggregate patient data, and one of mappings to descriptions of numeric values.

This later file was manually transformed into a long version for reading into our analysis environment.

```{r load_data}
PatientData <- read.csv("./Data/diabetic_data.csv", na.strings = "?")

Mappings <- read.csv("./Data/IDs_mapping_long.csv")

```

# Implimenting Exclusions

We initially decided to match our data to that of the original researchers. They reported 69984 observations after all exlcusion, but the public data set contained 101766. The below code attempts to match the researcher's count.

## Excluding Hospice and Expired Patients

```{r excuding_hospice_expired}
#Excluding Dispositions of Hospice and Expired
PatientData <- PatientData[PatientData$discharge_disposition_id %!in% c(13,14,11,20,21),]
```

Discharge of hospice (discharge ids 13 and 14) and Deceased (discharge ids 11, 20, 21) were excluded from our data, leaving `r nrow(PatientData)`

## Excluding Visits Beyond the First

```{r}

```




