---
title: "Group 3 - Chapter 9: Support Vector Machines Tutorial"
author: 
- Jolien Lam (j.lam@amc.uva.nl) – University of Amsterdam
- Yeng Moua (moua0199@umn.edu) – University of Minnesota
- Fortuna Vuthaj (fortuna.vuthaj@gmail.com) – University of Heilbronn/Heidelberg
- Ronald Buie (buierw@uw.edu) – University of Washington
- Jessica Kuo (U1010030142cmu.edu.tw) – Taipei Medical University
- Diane Walker (diane.walker@utah.edu) – University of Utah
output: html_notebook
---

```{r setup, include= FALSE}

```

This notebook documents work by group 3 utilizing support vector machines in the analysis of health data. Informaiton about the project can be found [here](https://iphie2018.wordpress.com/academic-program/).

# Front Matter

# Extraction

Data were acquired from ....

Two files are retrievable. One of the aggregate patient data, and one of mappings to descriptions of numeric values.

This later file was manually transformed into a long version for reading into our analysis environment.

```{r load_data}
PatientData <- read.csv("./Data/diabetic_data.csv", na.strings = "?")


Mappings <- read.csv("./Data/IDs_mapping_long.csv")
```